---
title: "Sensitivity and Specificity"
author: "Ai Sasho"
date: "October 5, 2015"
output: html_document
---


## Readings

***APM***

- ***Chapter 5 Measuring Performance in Regression Models*** (esp. ***5.2 The Variance Bias Trade-Off***)  (5 pages)
- ***Chapter 11 Measuring Performance in Classification Models*** (~20 pages)
- ***Chapter 7.4 K-Nearest Neighbors (regression)*** (2 pages)
- ***Chapter 13.5 K-Nearest Neighbors (classification)*** (3 pages)


```{r, echo=FALSE, results='hide', warning=FALSE }
packs <-  c('ggplot2', 'magrittr', 'dplyr', 'caret', 'AppliedPredictiveModeling')

for( nm in packs ) { 
  # message(nm)
  if( ! nm  %in% installed.packages()[,1]  ) install.packages(nm)
  library(nm, character.only = TRUE)
}

.. = NULL  # For Aesthetics

```


## EXERCISE 1: Resampling

`x` is a random variable. We want to not only know what the `mean(x)` is but want to calculate the uncertainty of `mean(x)`.  Measuring the uncertainty requires repeated measurements of `mean(x)`.

- Calculate the mean of `x`.
- Calculte the `sd( mean(x) )` using the **using 10-fold cross-validation**.  Create your own folds, show your work. (An example is for the Bootstrap is given as a hint. )


```{r}
set.seed(1) 
x <- runif(20,1,20)
x_mean = mean(x)

k=10

# CROSS-VALIDATION
# ... YOUR WORK HWEW

# split into 10 blocks hold out 1 block: take std of mean  

means_list <- c()
for (i in 1:k) {
  holdout_start_idx <- i + (i -1)
  holdout_end_idx <- holdout_start_idx + 1
  test <- x[c(-holdout_start_idx, -holdout_end_idx)]
  means_list <- c(means_list, mean(test))
}
sd_cv <- sd(means_list)


# BOOTSTRAP (EXAMPLE)
sd_boot <- sapply(1:k, function(i) sample(x,replace=TRUE) %>% mean ) %>% sd

```


- sd_cv   is: `r sd_cv`
- sd_boot is: `r sd_boot`



# Exercise 2: Binomial Metrics

Here's a really simple Model of Versicolor iris based on the **iris** data :

```{r}
set.seed(1)
data(iris)

qplot( data=iris, x=Petal.Length, y=Sepal.Length, color=Species )

# Create Dependent Variable
iris$Versicolor <- 
  ifelse( iris$Species == 'versicolor', "versicolor", "other" ) %>% as.factor
iris$Species = NULL 

wh <- sample.int( nrow(iris), size=nrow(iris)/2 ) #samples 1/2 num rows of iris
train <- iris[ wh,]
test <- iris[ -wh, ]


fit.glm <- glm( Versicolor ~ . - Sepal.Length, data=train, family=binomial )
```


Use the models to and write functions to calculate:

* Prevalence 
* Accuracy
* Error Rate / Misclassification Rate
* True Positive Rate  
* False Positive Rate
* True Negative Rate  
* False Negative Rate 
* Sensitivity 
* Specificity 
* Recall 
* Precision

The functions should take two logical vectors of the same length, `y` and `yhat`

```{r}

prevalence = function(y,yhat) {}
accuracy   =  .. 
error_rate = ..
tpr = ..
fpr = ..      # See Example
tnr = ..
sensitivity = ..
specificity = ..
recall = .. 
precision = ..

# EXAMPLE: fpr
# The FPR is THE NUMBER OF FALSE POSITIVES / NEGATIVES (TN+FP)

threshold = 0.5 
y = test$Versicolor == 'versicolor'
yhat = predict(fit.glm, test, type="response") > threshold

total_population <- sum(y) + sum(!y)

prevalence <- function(y, yhat) {
  # condition postives (= obsved positives) / total population
  sum(y) / total_population  
}
prevalence(y, yhat)

accuracy <- function(y, yhat) {
  (sum(yhat & y) + sum(!yhat & !y)) / total_population
}
accuracy(y, yhat)

error_rate <- function(y, yhat) {
  (sum(!yhat & y) + sum(yhat & !y)) / total_population
}
error_rate(y, yhat)

tpr <- function(y, yhat) {
  sum(yhat & y) / sum(y)
}
tpr(y, yhat)

fpr = function(y,yhat)
  sum(yhat & !y ) / # FP
  sum(! y)                # N

fpr(y,yhat)

tnr <- function(y, yhat) {
  sum(!yhat &  !y) / sum(!y)
}
tnr(y, yhat)

sensitivity <- function(y, yhat) {
  # TPR = num true positives / (num true positives + num false negatives)
  sum(yhat & y) / (sum(yhat & y) + sum(!yhat & y))
}
sensitivity(y, yhat)

specificity <- function(y, yhat) {
  # num true negatives / (num true negatives + num false positives)
  sum(!yhat & !y) / (sum(!yhat & !y) + sum(yhat & !y))
}
specificity(y, yhat)

recall <- function(y, yhat) {
  # same as sensitivity and TPR
  sum(yhat & y) / (sum(yhat & y) + sum(!yhat & y))
}
recall(y, yhat)

precision <- function(y, yhat) {
  # true positives / predicted positives
  sum(y) / sum(yhat)
}
precision(y, yhat)


```

- What is wrong with the modeling approach used?
- Answer: The true positive rate is low: 0.545, which means that the model could misidentify the true versicolr species as ther other species 
- Also, the error rate of 0.267 seems little high?




