---
title: "03-exercises"
author: "Ai Sasho"
date: "April 20, 2016"
output: html_document
---

## Readings

***APM***

- Chapter 4 "Over Fitting and Model Tuning"
- Chapter 12.2 "Logisitic Regression""


## Miscellaneous

I am still struggling with names ...

- Please send me your picture


## Assignment 

Note: The following will set-up your environment for this exercise. If you get an error stating that the packages have not been found, you need to install those packages.


```{r,echo=FALSE, warning=FALSE, message=FALSE}

packs <-  c('AppliedPredictiveModeling', 'ggplot2', 'magrittr', 'dplyr', 'caret', 'MASS')

for( nm in packs ) { 
  # message(nm)
  if( ! nm  %in% installed.packages()[,1]  ) install.packages(nm)
  library(nm, character.only = TRUE)
}

# Load data set into environment
data(FuelEconomy)     # See ?cars2010
fe <- dplyr::bind_rows(cars2010, cars2011, cars2012)    # Define Da


data("GermanCredit")  # see GermanCredit

... = NULL  # Needed for aesthetics 

```


## StepAIC


Using Fuel Economy data set from the **AppliedPredictiveModeling** Package.
- fit the simplest possible model using lm
- Use MASS::StepAIC to improve the model using forward stepwise regression
- Fit the "full" model using lm
- USe MASS::StepAIC to improve the model using backward stepwise regression 

```{r}

 # Your work here
 fit.min <- lm( FE ~ 1, fe )
 fit.min.forward <- stepAIC( fit.min, scope = ~ 1, direction = "forward")
 
 fit.full <- lm (FE ~ ., fe)
 fit.full.backward <- stepAIC( fit.full, scope = ~ 1, direction = "backward")

```

- Are they the same model? If not why?  Which is better? JUstify your answer.
They are different models, because fit.min only has one predicotr, whereas fit.full has all. 

```{r}
rmse <- function(y, yhat) {
  (y- yhat) ^2 %>% mean %>% sqrt
}

y.hat.fit.min.forward <- predict(fit.min.forward, fe)
rmse(fe$FE, y.hat.fit.min.forward)

y.hat.fit.full.backward <- predict(fit.full.backward, fe)
rmse(fe$FE, y.hat.fit.full.backward)
```

fit.full.backward is a better model, as when predicting the FE data, the RMSE is 3.437353, whereas, the RMSE for fit.min.forward is 8.096176. 

Is it a good assesment of the model? I used the training data to predict, and used the training data as y values and predicted as y-hat for RMSE. Probably I should be using a completely differnt set of data to asses the models?

## Logsitic and Inverse Logistic Transformation 

- Write an R function for the logistic function. The function should accept a `numeric` vector with values `[-Inf,Inf]` and produce a numeric vector in the the range `[0,1]`.

- Plot the logistic function from  `[-10,10]`

- Write a R function for the inverse logistic function. The function should accept a `numeric` vector with values `[0,1]` and prodcuce a numeric vector in the range `[-Inf,Inf]`

- Plot the Inverse Logistic function from `[0,1]`


**Hint:** For plotting curves see `?graphics::curve` or `?ggplot2::stat_function`


```{r}

logistic <- function(x) { 
  1/(1 + exp(-x))
}

qplot(x=c(-10:10), y =logistic(c(-10:10)), xlab="", ylab="logistic")

logistic_inv <- function(y) { 
  uniroot(logistic, c(-10,10), extendInt="yes")
}

#qplot(x=c(0, 1), y =logistic_inv(c(0.1, 1)), xlab="", ylab="inverse of logistic")

```

**NOTE"** These functions are quite handy, in evaluating logistic regression results. You may want to save these functions in your own package.  

```{r}
# DO NOT EDIT
c(-Inf,0,Inf) %>% logistic

c(0,0.5,1) %>% logistic_inv

```


## German Credit Model

Using the GermanCredit data from the **Caret** package/ UCI Machine Learning Library, create a model for `Class` ("Good" vs. "Bad" ). Show your model performance.  

```{r}
library(caret)
data(GermanCredit)
fit.glm <- glm(Class ~  ., data=GermanCredit, family="binomial")
fit.glm.both.directions <- stepAIC(fit.glm, data=GermanCredit, direction = "both")

summary(fit.glm.both.directions)

```



## Iterative Correlated Feature Removal 

- Implement Kuhn's iterative feature removal function described in **APM** Section 3.5, page 47

```{r}


feature_removal <- function(cor_matrix, threshold) {
 # find predictors that are highly correlated
 diag(cor_matrix) <- 0 # self correlations (1.0) won't be considered
 max_matrix <- which(cor_matrix == max(abs(cor_matrix)), arr.ind = TRUE)
 cols_corr_predictors <- max_matrix[,1] # columns of the strongly correlated predictors
 predictor1_col <- cols_corr_predictors[1]
 predictor2_col <- cols_corr_predictors[2]
 high_corr <- cor_matrix[predictor1_col, predictor2_col]
 
 # checks for stop condition
 if (high_corr > threshold) {
   # average correlation to the other predictors
   predictor1_mean <- mean(cor_matrix[predictor1_col])
   predictor2_mean <- mean(cor_matrix[predictor2_col])
   
   # determine the predictor with the higher correlation with the other predictors
   highest_corr_predictor <- predictor1_col
   if (predictor2_mean > predictor1_mean) {
     highest_corr_predictor <- predictor2_col
   }
  
   # recursive call until the stop condition meets
   col_row_remove <- -1 * highest_corr_predictor
   cor_matrix_new <- cor_matrix[col_row_remove, col_row_remove]
   feature_removal(cor_matrix_new, threshold)
 }
}

find_features <- function(x, threshold) {
  # calculate correlations
  cor_matrix <- cor(x)
  
  feature_removal(cor_matrix, threshold)
}
```

- Usage of above function:
find_features(iris[-5], threshold=0.9) 
- The function is not completely working yet..


## Synthetic Data (Optional)

Sometimes it is useful to "synthesize" feature data for to understand how a certain model behaves. 
Sythesize the following features 1000-element vectors: 

- x1: a normally distributed variable with `mean = 20` and standard deviation = 20 (`sd=8`).
- x2: a log-normally distributed feature with `meanlog = 1`, `sdlog=1.2`
- x3: a uniformly distributed feature with `min=0` and `max=50`. 

```{r}
nsamples = 20

x1 <- rnorm(nsamples,20,20)  
x2 <- rlnorm(nsamples, meanlog=1, sdlog = 1.2)
x3 <- runif(nsamples,0,50)

```

Next synthesis a response, `y` using the betas provided and an intercept that is normally distributed at 20 with standard deviation of 2. (**Hint:**  The betas thought of can be a vector or matrix)



```{r}

beta0 <- rnorm(nsamples,0,15)  # intercept!
beta1 <- 2.3
beta2 <- 4
beta3 <- 7

betas <- matrix( c(2.5, 4, 7), nrow=1  )  # 1x4 matrix

# x0 <- rep(1,nsamples) 

X  <- cbind(x1,x2,x3)  # 1000x4

y <- betas %*% t(X) %>% t
y <- y + beta0

qplot(y)
dat <- data.frame(y,X)

fit <- lm( y ~ . , dat )

coef(fit)

fit
```

- Did you recover the betas? 
- Is the model good?
- What happens if increase the value of `nsamples`? Decrease it?
- What transformations would you apply to x1? x2? x3? 

